Models of gravitational lens candidates from Space Warps CFHTLS

### 2017-09-06 parts/intro.tex
edited points flagged by IF

This paper follows up that study by applying SpaghettiLens to
candidates discovered through \SW.  We present results from
the modelling of 58 of the 59 lens candidates reported by
\cite{2016MNRAS.455.1191M}.  Each lens candidate was modelled
following a collaborative refinement process, where anyone interested
could improve the analysis by modifying an existing model or creating
a new one. Note the difference with respect to the main \SW 
project, where volunteers from a group of $\gtrsim10^4$ people make
independent contributions.  Each person is presented with a random
selection of survey-patches and invited to (in effect) vote on each.
The system estimates the skill level of each volunteer according to
test-patches interspersed with the real data, and weights their votes
accordingly \citep{2016MNRAS.455.1171M}.  There is an active forum for
volunteers, but since everyone is seeing different data samples with
minimal overlap, the forum has little if any influence on the votes.
\textcolor{red}{\bf IF: Please make sure the following is correct:}
In SpaghettiLens, the number of volunteers is significantly lower, but
the level of interaction is higher.  The resulting model represents a
consensus among contributors, as to the best that could be achieved
with the available data and software.

@@

\SW is a citizen science gravitational lens search
\citep{2016MNRAS.455.1171M}.  Its first run searched the CFHT Legacy
Survey, a survey carried out in five optical bands
($u^*$,$g^\prime$,$r^\prime$,$i^\prime$,$z^\prime$) covering a total
area of 160\,\sqdeg divided up into four fields W1 -- W4.  The
MegaPrime camera, used for the survey, has a field of view of
1\,\sqdeg, with 0.186$^{\prime\prime}$ pixels. The flux limit of the
survey is $g^\prime<$25.6AB (5\,$\sigma$) with a typical seeing
FWHM$\sim 0.7^{\prime\prime}$ \citep{2013MNRAS.433.2545E}. The cutouts
shown to the volunteers were colour composites of the $g^\prime$,
$r^\prime$ and $i^\prime$ bands from randomly chosen regions.
The programme \citep{2016MNRAS.455.1191M} resulted in 59 new lens candidates (of which 29 are
promising  \textcolor{red}{\bf IF: why are
  they promising? explain this here).}  Moreover, 82 previously known
lens candidates were ``rediscovered''.  The candidates have broadly
similar redshifts (0.2$<$z$<$1.0), Einstein radii between 0.7 and
5\,\arcsec, and arc magnitudes between 22 and 26 \textcolor{red}{\bf
  IF: in the $g^\prime$ band?}. These properties are similar to those
found by previous robotic searches like ArcFinder and RingFinder.

@@

A first version of \SpL\footnote{\textcolor{red}{\bf IF include web
    link: {\tt http://xxx.yyy.ch}}} was launched in May 2013, while
the first run of \SW was ongoing.  Fig.~\ref{fig:time} shows the
total number of models generated since launch for our lens candidates.
A small number of volunteers started creating models of possible lens
candidates from \SW that were debated in the discussion forum.  In the
beginning of August 2013 (around day 70 since release) a modelling
challenge for the candidate later to be identified as \sw{01} was
launched.  A group of five volunteers presented their results in an
online letter \textcolor{red}{\bf REFERENCE?} about 65 days later,
converging on a set of 30 models. In April 2015 (day 700),
the first major version of \SpL was released. At the same time,
the list of identified candidates was made available \citep[as a preprint
  of][]{2016MNRAS.455.1191M}.  At this stage, the volunteers were
asked to create models for all the lens candidates that were still
missing.  The work done for candidate \sw{29} show that the
volunteers converged on a favourable model much faster, after 15
models, generated in 30 days.  The total modelling effort resulted in
a set of 377 \SpL models for all but one of the 59 \SW lens
candidates; candidate \sw{39} was not modelled by any volunteer.  Over
its runtime, the system logged eight major contributors (defined as
volunteers that generated, at least, five models).

>>>>

This paper follows up that study by applying SpaghettiLens to
candidates discovered through \SW.  We present results from
the modelling of 58 of the 59 lens candidates reported by
\cite{2016MNRAS.455.1191M}.  Each lens candidate was modelled
following a collaborative refinement process, where anyone interested
could improve the analysis by modifying an existing model or creating
a new one. Note the difference with respect to the main \SW 
project, where volunteers from a group of $\gtrsim10^4$ people make
independent contributions.  Each person is presented with a random
selection of survey-patches and invited to (in effect) vote on each.
The system estimates the skill level of each volunteer according to
test-patches interspersed with the real data, and weights their votes
accordingly \citep{2016MNRAS.455.1171M}.  There is an active forum for
volunteers, but since everyone is seeing different data samples with
minimal overlap, the forum has little if any influence on the votes.
%\textcolor{red}{\bf IF: Please make sure the following is correct:}
% OK
In SpaghettiLens, the number of volunteers is significantly lower, but
the level of interaction is higher.  The resulting model represents a
consensus among contributors, as to the best that could be achieved
with the available data and software.

@@

\SW is a citizen science gravitational lens search
\citep{2016MNRAS.455.1171M}.  Its first run searched the CFHT Legacy
Survey, a survey carried out in five optical bands
($u^*$,$g^\prime$,$r^\prime$,$i^\prime$,$z^\prime$) covering a total
area of 160\,\sqdeg divided up into four fields W1 -- W4.  The
MegaPrime camera, used for the survey, has a field of view of
1\,\sqdeg, with 0.186$^{\prime\prime}$ pixels. The flux limit of the
survey is $g^\prime<$25.6AB (5\,$\sigma$) with a typical seeing
FWHM$\sim 0.7^{\prime\prime}$ \citep{2013MNRAS.433.2545E}. The cutouts
shown to the volunteers were colour composites of the $g^\prime$,
$r^\prime$ and $i^\prime$ bands from randomly chosen regions.
The programme resulted in 59 new lens candidates, of which 29 are
considered high quality \citep{2016MNRAS.455.1191M}.
% \textcolor{red}{\bf IF: why are they promising? explain this here).}
Moreover, 82 previously known lens candidates were ``rediscovered''.
The candidates have broadly similar redshifts ($0.2<z<1.0$), Einstein
radii between 0.7 and 5\,\arcsec, and arc magnitudes between 22 and 26
\textcolor{red}{\bf IF: in the $g^\prime$ band?}. These properties are
similar to those found by previous robotic searches like ArcFinder and
RingFinder.

@@

A first version of \SpL
% \footnote{\textcolor{red}{\bf IF include web
%  link: {\tt http://xxx.yyy.ch}}}
% Not currently online
was launched in May 2013, while
the first run of \SW was ongoing.  Fig.~\ref{fig:time} shows the
total number of models generated since launch for our lens candidates.
A small number of volunteers started creating models of possible lens
candidates from \SW that were debated in the discussion forum.  In the
beginning of August 2013 (around day 70 since release) a modelling
challenge for the candidate later to be identified as \sw{01} was
launched.  A group of five volunteers presented their results in an
online letter
% \textcolor{red}{\bf REFERENCE?} about 65 days later,
% Not available any more.
converging on a set of 30 models. In April 2015 (day 700),
the first major version of \SpL was released. At the same time,
the list of identified candidates was made available \citep[as a preprint
  of][]{2016MNRAS.455.1191M}.  At this stage, the volunteers were
asked to create models for all the lens candidates that were still
missing.  The work done for candidate \sw{29} show that the
volunteers converged on a favourable model much faster, after 15
models, generated in 30 days.  The total modelling effort resulted in
a set of 377 \SpL models for all but one of the 59 \SW lens
candidates; candidate \sw{39} was not modelled by any volunteer.  Over
its runtime, the system logged eight major contributors (defined as
volunteers that generated, at least, five models).

### 2017-09-06 parts/massmodel.tex
edited points flagged by IF

Once a spaghetti diagram has been drawn on a web browser, it is
forwarded to a server-side numerical framework, which searches for
mass maps consistent with the image locations, parities and time
ordering, given by the modeller.  The mass maps are made up of mass
tiles and are free-form, but are required to be concentrated around
the identified lens centre \citep[see][for the precise formulation of
  the search problem]{2014MNRAS.445.2181C}. Assuming such mass
distributions can be found -- which in practice is usually the case --
a statistical ensemble of two-dimensional mass maps is returned.  This
ensemble, along with derived quantities and uncertainties, makes up
one SpaghettiLens model.

The mass will naturally depend on the lens and source redshifts, which
are unknown when a lens candidate is first identified.  However, this
is not a problem, because a model can be trivially rescaled to use
better redshift values, as they become available.  This work applies
photometric redshifts from the CFHTLS pipeline
\citep{2009A&A...500..981C} to the candidate lensing galaxies; the
values range from z=0.2 to 1 (see Table~\ref{tab:models}).
These redshifts entail rather big uncertainties, up to a few ten
percent.  Since, however, the angular critical density is proportional
to $d_l d_s d_{ls}^{-1}$ \citep[consult Equation~A1
  in][]{2015MNRAS.447.2170K} it is negligible for this tentative
analysis. \textcolor{red}{\bf IF: I would not say this is
  ``negligible'' but, rather, that the error in the redshift would lead to
  surface mass density uncertainties of XX\%, which, for the
  aim of the tentative analysis presented here, is an acceptable
  result. You should also check and quantify properly the
  uncertainty in the photo-z. What is the value claimed by the
  CFHTLS group? I would say 5-10\% is a decent level, but you
are mentioning 10s of \% which is rather extreme!}.
As the mass normalisation depends only weakly on the source redshift,
provided it is higher than the lens redshift, we adopted a redshift
z=2 for all sources, unless an unambiguous photometric redshift is
available. \textcolor{red}{\bf IF: Again, a simple quantification of the
change you expect, for instance, if the source is at z=3 would be good.}

@@

Fig.~\ref{fig:encl} shows the enclosed-mass profiles, expressed as the
average convergence, $\kappa$, within circles of given projected
radius.  The uncertainties \textcolor{red}{\bf IF: at the BLAH level}
are included.  Appendix~\ref{subsec:hires} describes the improvements made
since our earlier work \citep{2015MNRAS.447.2170K}, to allow for
steeper profiles in the inner regions.  The enclosed mass is typically
best constrained at the notional Einstein radius, becoming more
uncertain at larger and smaller radii.

>>>>

Once a spaghetti diagram has been drawn on a web browser, it is
forwarded to a server-side numerical framework, which searches for
mass maps consistent with the image locations, parities and time
ordering, given by the modeller.  The mass maps are made up of mass
tiles and are free-form, but are required to be concentrated around
the identified lens centre \citep[see][for the precise formulation of
  the search problem]{2014MNRAS.445.2181C}.  The modeller can also set
various options for the search, such as the number of mass tiles and
the extent of the mass map; all options have defaults.  Typically,
there are $\sim500$ mass tiles arranged in a disc, centred on the
lensing galaxy and extending to twice the radius of the outermost
image.  By default, the mass distributed is required to have
$180^\circ$ rotation symmetry, but this option can be unset.  Assuming
mass distributions can be found -- which in practice is usually the
case -- a statistical ensemble of two-dimensional mass maps is
returned.  This ensemble, along with derived quantities and
uncertainties, makes up one SpaghettiLens model.

The mass will naturally depend on the lens and source redshifts, which
are unknown when a lens candidate is first identified.  However, this
is not a problem, because a model can be trivially rescaled to use
better redshift values, as they become available.  The mass
normalisation of the models is proportional, through the angular
critical density, to the factor $d_L\, d_S\, d_{LS}^{-1}$, where
$d_L,d_S$ and $d_{LS}$ are the usual angular-diameter distances.  In
the redshift ranges typical of galaxy lensing, the normalisation
factor is roughly proportional to the lens redshift, and weakly
dependent on the source redshift.  This work applies photometric
redshifts from the CFHTLS pipeline \citep{2009A&A...500..981C} to the
candidate lensing galaxies; the values range from $z=0.2$ to 1 (see
Table~\ref{tab:models}).  The source redshifts are assumed as $z=2$,
unless an unambiguous photometric redshift is available.  The lens
redshifts redshifts entail rather big uncertainties, up to a few tens
of percent.  The lensing masses are uncertain at the same level.  On
the other hand, the lensing masses in the sample range from
$10^{11}M_\odot$ to $10^{13}M_\odot$, hence the redshift uncertainty
is not so important in this preliminary analysis.
% \textcolor{red}{\bf IF: I would not say this is
% ``negligible'' but, rather, that the error in the redshift would lead to
%  surface mass density uncertainties of XX\%, which, for the
%  aim of the tentative analysis presented here, is an acceptable
%  result. You should also check and quantify properly the
%  uncertainty in the photo-z. What is the value claimed by the
%  CFHTLS group? I would say 5-10\% is a decent level, but you
%  are mentioning 10s of \% which is rather extreme!}.
% \textcolor{red}{\bf IF: Again, a simple quantification of the
% change you expect, for instance, if the source is at z=3 would be good.}

@@

Fig.~\ref{fig:encl} shows the enclosed-mass profiles, expressed as the
average convergence, $\kappa$, within circles of given projected
radius.  Uncertainties are included (see the figure caption for
details).
% \textcolor{red}{\bf IF: at the BLAH level}
% Explained in the caption.
Appendix~\ref{subsec:hires} describes the improvements made
since our earlier work \citep{2015MNRAS.447.2170K}, to allow for
steeper profiles in the inner regions.  The enclosed mass is typically
best constrained at the notional Einstein radius, becoming more
uncertain at larger and smaller radii.
